A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-29 paper/draft.tex
appointment

Empirical research on allocation of public sector jobs has consistently found that politically motivated allocation of public sector jobs is an important determinant of public employment in the developed and developing world \citep{finan2017personnel}. In the United States, appointments to the federal bureaucracy involve considerations of party loyalty and ideological alignment with the president \citep{lewis2010politics, hollibaugh2014presidents}. In Brazil, whether it be at the presidential \citep{pracca2011political} or at the municipal level \citep{colonnelli2018patronage,brollo2017victor}, politicians have discretion over bureaucratic appointments and frequently use these to further their political goals.

A set of explanations have been proposed to explain the determinants of this allocation, whether it be to reduce frictions in policy implementation due to ideological divergence \citep{krause2016experiential}, exploit the benefits of strong ties \citep{toral2019benefits} or to reward followers for campaign contributions \citep{colonnelli2018patronage}. These findings have provided important insights into the drivers of patronage appointments into employment in the public sector. Yet as important as the drivers of public sector appointment are the characteristics of those entering the bureaucracy \citep{finan2015personnel}. Who becomes a bureaucrat? Are political appointees systematically different from their non-partisan counterparts? And where, within the bureaucracy, do these patronage appointees go?

>>>>

Empirical research on allocation of public sector jobs has consistently found that political motivations are an important determinant of public employment in the developed and developing world \citep{finan2017personnel}. In the United States, appointments to the federal bureaucracy involve considerations of party loyalty and ideological alignment with the president \citep{lewis2010politics, hollibaugh2014presidents}. In Brazil, whether it be at the presidential \citep{pracca2011political} or at the municipal level \citep{colonnelli2018patronage,brollo2017victor}, politicians have discretion over bureaucratic appointments and frequently use these to further their political goals.

A set of explanations have been proposed to explain the determinants of this allocation, whether it be to reduce frictions in policy implementation due to ideological divergence \citep{krause2016experiential}, exploit the benefits of strong ties \citep{toral2019benefits} or to reward followers for campaign contributions \citep{colonnelli2018patronage}. These findings have provided important insights into the drivers of patronage appointments into employment in the public sector. Yet the literature has been relatively silent about the characteristics of those entering the bureaucracy \citep{finan2015personnel}. Who becomes a bureaucrat? Are political appointees systematically different from their non-partisan counterparts? And where, within the bureaucracy, do these patronage appointees go?
